Dark matter haloes: a multistream view

### 2018-06-15 Draft/Chapter5/Paper2017b_Nobmath.tex
conclusion updated

Identifying the haloes from the distribution of particles in N-body simulations is one of the problems attracting both considerable interest and efforts. In this chapter, we propose a novel framework for detecting potential dark matter haloes using the field unique for dark matter  -- multistream field. The current status of halo finding techniques are summarized in Section \ref{sec:intro}. A review of the DM particle clustering in a one-dimensional dimensional universe is made in Section \ref{sec:HaloFormation}, and the concept of multistream field is extended to higher dimensions. The multistream field is computed on the cosmological simulations described in Section \ref{sec:ch5_simulation}. The halo identification framework in this field is described in Section \ref{sec:haloDetection}. This algorithm isolates convex regions of the multistream field using Hessian eigenvalues, each enclosing a local multistream maximum. Without employing any non-local thresholds that several halo finders generally use, these convex multistream regions are identified as potential halo sites. We also illustrate the significance of multistream refinement and softening scales in finding sub-haloes. However, this chapter does not focus on adaptive multi-scale analyses for substructure studies. A few properties of the multistream haloes are discussed in Section \ref{sub:HaloProperties}, and comparison of these haloes with AHF and FOF algorithms is done in Section \ref{sub:compareHalo}. We also discuss the spatial distribution of the dark matter haloes along the percolating web structure.

\section{Overview of halo finders} 
\label{sec:intro}

@@

Multistream field $n_{str}(\mathbf{x})$ is calculated on the GADGET-2 snapshots at $z=0$ using the tessellation scheme by \cite{Shandarin2012}. The multistream field can be computed at the native resolution of the Lagrangian grid of the simulation, i.e., at refinement factor of $l_l/l_d = 1$ (where $l_l$ is the inter-particle separation in Lagrangian grid and $l_d$ is the side length resolution of diagnostic grid). Arbitrarily high refinement factors can be utilized in computing multistream fields as well, for example  $l_l/l_d = 8$ for the halo multistream environment shown in \autoref{fig:full}. For analysis of full simulation boxes, we restrict $l_l/l_d$ to $1$ and $2$.

@@

\subsection{Halo finder algorithm}
\label{subsec:technique}

@@

However, the $\lambda_3$-halo boundary is different from any constant multistream contour. That is, from the function $n_{str}(\mathbf{x})$, convex regions in the four-dimensional function space $(-n_{str}, x, y, z)$ are projected onto three-dimensional co-ordinate space using eigenvalues. This is different from projecting `iso-multistream' slice onto three-dimensional co-ordinate space. Appendix \ref{appendix:Eigen} illustrates the difference in the two approaches for a one-dimensional function.

@@

Haloes from the three finders in \autoref{fig:FinderCompare3} 
also show other differences in the halo boundaries. While all the AHF-haloes are spherical by definition, the FOF-haloes are irregular. The boundaries of the $\lambda_3$ haloes are not spherical either, but are more compact than FOF-haloes and in addition they are convex surfaces by design. At some junctions of the filaments, FOF identifies a large region as belonging to single halo, whereas AHF and our algorithm detect multiple isolated haloes. Each isolated $\lambda_3$-halo region enclose one maximum of multistream field, thus guaranteeing that multiple haloes are always resolved. On the other hand, a linking length cut-off or a constant threshold on scalar fields may enclose regions with multiple local maxima ( For one-dimensional fields, an illustration of this is shown in Appendix \ref{appendix:Eigen}).

@@

\section{Discussion}
\label{sec:discussion}

@@

\section{Summary}
\label{sec:summary}

>>>>

Identifying the haloes from the distribution of particles in N-body simulations is one of the problems attracting both considerable interest and efforts. In this chapter, we propose a novel framework for detecting potential dark matter haloes using the field unique for dark matter  -- multistream field. The current status of halo finding techniques are summarized in Section \ref{sec:5intro}. A review of the DM particle clustering in a one-dimensional dimensional universe is made in Section \ref{sec:HaloFormation}, and the concept of multistream field is extended to higher dimensions. The multistream field is computed on the cosmological simulations described in Section \ref{sec:ch5_simulation}. The halo identification framework in this field is described in Section \ref{sec:haloDetection}. This algorithm isolates convex regions of the multistream field using Hessian eigenvalues, each enclosing a local multistream maximum. Without employing any non-local thresholds that several halo finders generally use, these convex multistream regions are identified as potential halo sites. We also illustrate the significance of multistream refinement and softening scales in finding sub-haloes. However, this chapter does not focus on adaptive multi-scale analyses for substructure studies. A few properties of the multistream haloes are discussed in Section \ref{sub:HaloProperties}, and comparison of these haloes with AHF and FOF algorithms is done in Section \ref{sub:compareHalo}. We also discuss the spatial distribution of the dark matter haloes along the percolating web structure.

\section{Overview of halo finders} 
\label{sec:5intro}

@@

Multistream field $n_{str}(\mathbf{x})$ is calculated on the GADGET-2 snapshots at $z=0$ using the tessellation scheme by \cite{Shandarin2012}. The multistream field can be computed at the native resolution of the Lagrangian grid of the simulation, i.e., at refinement factor of $l_l/l_d = 1$ (where $l_l$ is the inter-particle separation in Lagrangian grid and $l_d$ is the side length resolution of diagnostic grid). Arbitrarily high refinement factors can be utilized in computing multistream fields as well, for example  $l_l/l_d = 8$ for the halo multistream environment shown in \autoref{fig:3full}. For analysis of full simulation boxes, we restrict $l_l/l_d$ to $1$ and $2$.

@@

\subsection{Halo finder algorithm}
\label{subsec:5technique}

@@

However, the $\lambda_3$-halo boundary is different from any constant multistream contour. That is, from the function $n_{str}(\mathbf{x})$, convex regions in the four-dimensional function space $(-n_{str}, x, y, z)$ are projected onto three-dimensional co-ordinate space using eigenvalues. This is different from projecting `iso-multistream' slice onto three-dimensional co-ordinate space. Appendix \ref{appendix:bEigen} illustrates the difference in the two approaches for a one-dimensional function.

@@

Haloes from the three finders in \autoref{fig:FinderCompare3} 
also show other differences in the halo boundaries. While all the AHF-haloes are spherical by definition, the FOF-haloes are irregular. The boundaries of the $\lambda_3$ haloes are not spherical either, but are more compact than FOF-haloes and in addition they are convex surfaces by design. At some junctions of the filaments, FOF identifies a large region as belonging to single halo, whereas AHF and our algorithm detect multiple isolated haloes. Each isolated $\lambda_3$-halo region enclose one maximum of multistream field, thus guaranteeing that multiple haloes are always resolved. On the other hand, a linking length cut-off or a constant threshold on scalar fields may enclose regions with multiple local maxima ( For one-dimensional fields, an illustration of this is shown in Appendix \ref{appendix:bEigen}).

@@

\section{Discussion}
\label{sec:5discussion}

@@

\section{Summary}
\label{sec:5summary}
